### 2020-07-06 p5/proposal.tex
Initial commit.

\documentclass{article}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{bm}
\begin{document}
\title{Verifying research using ODEs in NLP}
\maketitle
The difference equation to be examined is
\begin{equation}\label{e1}
h_t = (1 - z_{t-1})\odot h_{t-1} + z_t \odot \hat{h}_t,
\end{equation}
where
\begin{eqnarray}
\hat{h}_t &=& \tanh(W \odot (r_t \odot h_{t - 1} + x_t) \label{e2} \\
r_t &=& \sigma\left(w^l x_t + U^l h_{t - 1}\right) \label{e3} \\
z_t &=& \sigma\left(w^z x_t + U^z h_{t - 1}\right) \label{e4}
\end{eqnarray}
Now suppose I want to investigate equation \eqref{e1} and I start writing a
program to do so. It is clear that I need the following:
\begin{enumerate}
\item $h_0$, the initial value. Now thankfully, this difference equation
is a discrete version of a differential equation. We hope that the ODE is
stable and therefore where we start does not make a difference. (Actually
it does make a difference when $h$ has more than one dimension. But for
the moment, let's gloss it over.) Let us choose $h_0 = 0$.
\item $z_0$. We cannot use equation \eqref{e4} to proceed because it will
then demand $h_{-1}$ from us and there is no $h$ before `initial time'. Let
us make another assumption $z_0 = 0$ to permit us to proceed.
\item $x_t$, there is nothing that tells us what this is.
\item $w^l, w^z, U^l, U^z, W$, we need to know what these objects are before
we proceed.
\end{enumerate}
You now see that we have $5$ unknowns $w^l, w^z, U^l, U^z, W$. We do not 
even know whether they are vectors or matrices, constants or variables. 
More seriously, we do not know that $x_t$ is. We need it for all $t \ge 0$.

I think we are not in a stage where we can conduct a mathematical analysis
of difference equations. We must define our objects precisely before we
proceed.

\end{document}