### 2020-11-29 optim/amber-india/prob3.tex
The third problem.

\documentclass{article}
\begin{document}
\begin{enumerate}
\item
Let $s$ be the selling price, $c$ be the cost price, $u$ be the units in stock
and $D$ be the random demand. Then the profit is
\begin{equation}
p = s \min(u, D) - cu.
\end{equation}

\item The decision variable is $u$, the units stocked.

\item The optimal number of units is
\begin{equation}
u = F^{-1}\left(\frac{s - c}{s}\right),
\end{equation}
where $F$ is the cumulative distribution function of the random demand.

In the problem $p = 12$ dollars. The cost per glass is $75/25 = 3$ 
dollars. Therefore, 
\begin{equation}
u = F^{-1}\left(\frac{12 - 3}{12}\right) = F^{-1}(0.75).
\end{equation}
The demand $D \sim U(200, 500)$ so that $u = 200 + 300 \times 0.75 = 425$.

Alternatively you can use the R function 
\begin{verbatim}
u <- qunif(0.75, 200, 500)
\end{verbatim}
to get the same answer.

The optimal profit is $(12 - 3) \times 425/25 = 153$ dollars.
\end{enumerate}
\end{document}